---
title: "A Visualization of U.S. Couples' Money Arrangements"
author: "Joanna R. Pepin"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
      out_dir <- "docs";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_file='index.html',
                        output_dir=file.path(dirname(inputFile), out_dir))})
output:
   html_document:
    df_print: paged
bibliography: references.json
---

# Abstract
Using newly available nationally representative U.S. data from the [National Couples' Health and Time Study](https://www.icpsr.umich.edu/web/DSDR/studies/38417), this visualization presents 2020–2021 estimates of couples' (n=3,396) money arrangements by marital and parental status, age group, and for different-gender and same-gender couples. It shows cohabiting couples continue to be much less likely than married couples to pool all their money, and the gap is apparent across key demographic characteristics. Among married couples, 66% of different-gender couples put all their money together, compared with 45% of women partnered with women and 50% of men partnered with men. Young couples are as likely as older couples to put all money together when legally married.
  
  
### Keywords  
money, couples, marriage, cohabitation, visualization  
  
  
--- 
  
A recent survey experiment showed the most commonly selected money arrangement for a fictional couple was to combine some of their money and keep some money separate  (@pepin_beliefs_2019). Whether U.S. couples’ behaviors have changed in ways consistent with these public beliefs has remained an open empirical question because the most recent publicly available data on U.S. couples’ money arrangements comes from the 2012 International Social Survey Programme Family and Changing Gender Roles module. This 10-year gap has hindered crucial knowledge about societal trends in the pursuit of autonomy vs. collectivism in relationships. Couples’ treatment of money also has important substantive implications regarding relationship quality and persistent gender inequality within families. Moreover, a comparison between different-gender and same-gender couples’ money arrangements has been unobtainable using nationally representative survey data until now.  
  
  
Figure 1 presents updated estimates of couples’ approaches to money using 2020–2021 data from the National Couples’ Health and Time Study (NCHAT), a nationally representative sample of married and cohabiting couples living in the United States (@kamp_dush_national_2022). NCHAT oversampled sexual and gender diverse families, and is nationally representative of married and cohabiting couples (aged 20–60) living in the United States. Couples’ money arrangements are presented by marital and parental status, for different-gender and same-gender couples, and by age group. Results are disaggregated by legal marital status, which has consistently been found to be the determining characteristic influencing couples’ integration of money (@eickmeyer_whats_2019). Each panel in the figure contains a waffle chart consisting of 100 smaller squares colored according to the proportion of couples who report they (1) put all of their money together, (2) put some money together, or (3) keep all of their money separate. Additional details are available in the [online supplement](https://jrpepin.github.io/NCHAT_Money/Supplementary-Material.html) and data access and replication code are available at:  [https://github.com/jrpepin/NCHAT_Money](https://github.com/jrpepin/NCHAT_Money).  
  
Differences by marital status are apparent, with 66% of married couples reporting they pool everything compared with 23% of cohabiting couples taking a complete pooling approach (p<.05). The marriage-cohabitation gap is visible across sub-groups and chi-squared tests confirm it is statistically significantly different within each demographic characteristic (p<.05). Married couples who are parenting are more likely than married couples who are not parenting to put all their money together (p<.05). Two-thirds of married different-gender couples put all of their money together compared with 50% of married men partnered with men and 45% of married women partnered with women (p<.05). Cohabiting couples’ approaches to money are relatively consistent within sub-groups and are not statistically significantly different.    
  
  
```{r setup, echo = FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

## WARNING: Remove the leading # to install packages below first time. 
## Change filepaths below first time

  # if (!require(remotes)) install.packages("remotes") 
  # if (!require(devtools)) install.packages("devtools") 
  # remotes::install_github("fsolt/icpsrdata")                      # download ICPSR data
  # devtools::install_git("https://git.rud.is/hrbrmstr/waffle.git") # install waffle package not on CRAN
  # install.packages("pacman")                                      # Install pacman package if not installed
  # remotes::install_github("paleolimbot/rbbt")                     # for Zotero citations in RStudio

# Installs and loads packages automatically
library("pacman")                  # Load pacman package

# Install packages not yet installed & load them
pacman::p_load(
       here,       # relative file paths
       foreign,    # read data
       dplyr,      # variable processing
       tidyr,      # reshaping data
       forcats,    # reverse factor variables
       srvyr,      # calc % with survey weights
       MESS,       # round prop & preserve sum to 100%
       gtsummary,  # pretty weighted tables
       ggplot2,    # graphing
       ggtext,     # automatically wraps the text inside figure
       colorspace) # color palettes  

library("waffle")                 
library("icpsrdata")

```
  
  
```{r filepaths, echo = FALSE}
# Set the project directory to the current working directory.
# Change the filepath to where the data was downloaded.
projDir <- here::here()             # File path to this project's directory
dataDir <- "./../../Data/NCHAT"     # File path to where data will be downloaded

```
  
  
```{r data-1, echo=FALSE, eval = FALSE}
# The icpsr_download function will ask for the login information required by ICPSR: the user’s email and password.
## URL for manual download: https://www.icpsr.umich.edu/web/DSDR/studies/38417
icpsr_download( file_id = 38417,
                download_dir = file.path(dataDir))
```
  
  
```{r data-2, echo=FALSE, results= 'hide'}
# Load the data and create a new dataframe containing only the variables of interest.  

load(file=file.path(dataDir, "ICPSR_38417/DS0001", '38417-0001-Data.rda'))

data <- da38417.0001 %>%
  select(RESPONSEID, WEIGHT_MAINRESPONDENT, 
         Q40,       # money treatment
         MARITAL_1, # legal marital status
         C18D_B,    # parent identifier using skip logic
                    # {ends_with("HHR16"), starts_with("A1_HHR17_"), starts_with("A1_HHR18_") is masked by ICPSR]}
         D2, HHR5,  # couple type
         AGE_SD)

```
  
  
```{r vars, echo=FALSE}
### creates a function so level order remains the same as argument order
fct_case_when <- function(...) {
  args <- as.list(match.call())
  levels <- sapply(args[-1], function(f) f[[3]])  # extract RHS of formula
  levels <- levels[!is.na(levels)]
  factor(dplyr::case_when(...), levels=levels)
}

## Variables
data <- data %>%
  mutate(
    # money treatment
    money = fct_case_when(
      Q40 == "(3) Put all of our money together"     ~ "Put all money together",
      Q40 == "(2) Put some money together"           ~ "Put some money together",
      Q40 == "(1) We keep all of our money separate" ~ "Keep all money separate"),
    # marital status
    marco = fct_case_when(
      MARITAL_1  == "(1) Legal marriage"             ~ "Married",
      TRUE                                           ~ "Cohabiting"),
    # parental status
    parent = fct_case_when(
      !is.na(C18D_B)                                 ~ "Parenting",
      TRUE                                           ~ "Not parenting"),
    # couple type
    couple = fct_case_when(
     (D2 == "(1) Man"   & HHR5 == "(2) Woman")       |
     (D2 == "(2) Woman" & HHR5 == "(1) Man")         ~ "Different-gender",
      D2 == "(2) Woman" & HHR5 == "(2) Woman"        ~ "Woman-woman",
      D2 == "(1) Man"   & HHR5 == "(1) Man"          ~ "Man-man"),
    # age
    age = fct_case_when(
      AGE_SD == "(1) 18-34"                          ~ "Age: 20-34",      # change bottom code to match codebook
      AGE_SD == "(2) 35-50"                          ~ "Age: 35-50",
      AGE_SD == "(3) 51-69"                          ~ "Age: 51-60")) %>% # change top code to match codebook
  select(money, marco, parent, couple, age, WEIGHT_MAINRESPONDENT, RESPONSEID) %>%
  drop_na() # drop respondents with any missing data

```
  
  
```{r survey, echo=FALSE, results='hide'}

waffle_svy <- data %>%
  as_survey_design(id = RESPONSEID,
                   weights = WEIGHT_MAINRESPONDENT)

waffle_svy %>%
  select(c(-RESPONSEID, -WEIGHT_MAINRESPONDENT)) %>%
  tbl_svysummary(statistic = all_categorical() ~ "{n_unweighted} ({p}%)",
                 label = list(money    ~ "Money Treatment",
                              marco    ~ "Marital Status",
                              parent   ~ "Parental Status",
                              couple   ~ "Couple Type",
                              age      ~ "Age")) %>%
  modify_header(update = stat_0 ~ "**N = {N_unweighted}**")
 
```
<p>  
  
  
  
```{r waffle-1, echo=FALSE, results='hide'}

# all -- just for reference
df0 <- waffle_svy %>%
  group_by( money) %>%
  # create summary proportions
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>% 
  # round with preserving to 100%
  mutate(vals     = MESS::round_percent(vals), 
         vals_low = MESS::round_percent(vals_low), 
         vals_upp = MESS::round_percent(vals_upp))

# marital status
df1 <- waffle_svy %>%
  group_by(marco, money) %>%
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>% 
  mutate(vals     = MESS::round_percent(vals), 
         vals_low = MESS::round_percent(vals_low), 
         vals_upp = MESS::round_percent(vals_upp), 
         group    = "parent",
         subgroup = "All")

# parents
df2 <- waffle_svy %>%
  group_by(marco,parent, money) %>%
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>%
  mutate(vals     = MESS::round_percent(vals), 
         vals_low = MESS::round_percent(vals_low), 
         vals_upp = MESS::round_percent(vals_upp), 
         group    = "parent") %>%
  rename(subgroup = parent)

# couple type
df3 <- waffle_svy %>%
  group_by(marco,couple, money) %>%
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>%
  mutate(vals     = MESS::round_percent(vals), 
         vals_low = if_else(vals_low < 0, 0, vals_low), # round negative value to zero
         vals_low = MESS::round_percent(vals_low), 
         vals_upp = MESS::round_percent(vals_upp), 
         group    = "couple") %>%
  rename(subgroup = couple)

# age groups
df4 <- waffle_svy %>%
  group_by(marco,age, money) %>%
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>%
  mutate(vals     = MESS::round_percent(vals), 
         vals_low = MESS::round_percent(vals_low), 
         vals_upp = MESS::round_percent(vals_upp), 
         group    = "age") %>%
  rename(subgroup = age)

# Combine datasets
waffle_data <- df1 %>%
  full_join(df2) %>%
  full_join(df3) %>%
  full_join(df4) %>%
    mutate(group = as_factor(group),
    subgroup = as_factor(subgroup))

head(waffle_data, n = 10) # to check results

```
  
```{r waffle-2, echo=FALSE, results='hide'}
# Prep the data for plotting

## Create a combo variable because geom_waffle doesn't work with facet_grid()
waffle_data$c<-paste0(as.character(waffle_data$marco),"\n", as.character(waffle_data$subgroup))

waffle_data$c <- factor(waffle_data$c, 
                              levels = c("Married\nAll", "Married\nParenting", "Married\nNot parenting",
                                         "Cohabiting\nAll", "Cohabiting\nParenting", "Cohabiting\nNot parenting",
                                         "Married\nDifferent-gender", "Married\nWoman-woman", "Married\nMan-man",
                                         "Cohabiting\nDifferent-gender", "Cohabiting\nWoman-woman", "Cohabiting\nMan-man",
                                         "Married\nAge: 20-34", "Married\nAge: 35-50", "Married\nAge: 51-60",
                                         "Cohabiting\nAge: 20-34", "Cohabiting\nAge: 35-50", "Cohabiting\nAge: 51-60"), 
                              ordered = FALSE)

## Create a variable for plotting the values & correct top rows
waffle_data$yaxis <- round(waffle_data$vals/10) 
waffle_data$yaxis[waffle_data$c == "Cohabiting\nParenting" & 
                  waffle_data$money == "Keep all money separate"] <- 4 # correction re: rounded off chart
waffle_data$yaxis[waffle_data$c == "Cohabiting\nNot parenting" & 
                  waffle_data$money == "Keep all money separate"] <- 6 # move up a row 
waffle_data$yaxis[waffle_data$c == "Married\nWoman-woman" & 
                  waffle_data$money == "Keep all money separate"] <- 3 # move up a row 
waffle_data$yaxis[waffle_data$c == "Cohabiting\nDifferent-gender" & 
                  waffle_data$money == "Keep all money separate"] <- 6 # move up a row 
waffle_data$yaxis[waffle_data$c == "Cohabiting\nAge: 20-34" & 
                  waffle_data$money == "Keep all money separate"] <- 6 # move up a row 
waffle_data$yaxis[waffle_data$c == "Cohabiting\nAge: 51-60" & 
                  waffle_data$money == "Keep all money separate"] <- 6 # move up a row 

```
  
```{r waffle-3, echo = FALSE, fig.align='center', fig.width=6.5, fig.height=6.5}

p1<- waffle_data %>%
  ggplot(aes(fill = money, values = vals)) +
  geom_waffle(color = "white", size = .2, n_rows = 10, flip = TRUE) +
  geom_text(aes(x=1.1, y=yaxis, label=vals), 
            color    = "black", 
            size     = 3,
            fontface ="bold",
            position = position_stack(reverse = TRUE)) +
  facet_wrap(vars(c), ncol = 6) +
  scale_fill_manual(values = sequential_hcl(3, palette = "OrYel")) +
  coord_fixed(ratio = 1) +
  #  guides(fill = guide_legend(reverse = TRUE)) + # fct_rev()
  theme_minimal() +
  theme(panel.grid      = element_blank(),
        axis.title      = element_blank(),
        axis.text.x     = element_blank(),
        axis.text.y     = element_blank(),
        legend.title    = element_blank(),
        plot.title.position   = "plot",
        plot.caption.position =  "plot",
        plot.caption          = element_textbox_simple(),
        panel.spacing         = unit(0, "lines"),
        plot.subtitle         = element_text(face = "italic"),
        legend.position       = "top") +
  labs(title = "U.S. couples' money arrangements by demographic groups (n = 3,396)",
       subtitle = "Percentage of U.S. couples (aged 20-60) who selected _____________",
       caption = "**Figure 1.** Data are from the National Couples' Health and Time Study (U.S.), 2020-2021. 
       Sampling weights are used to adjust for the complex sampling process of the survey, 
       such as the oversample of sexual and gender diverse families.
       Chi-squared tests confirm all demographic group comparisons between married and cohabiting couples are statistically significant (p<.05).
       Among married couples, parents and different-gender couples are more likely than their counterparts to report 
       they 'put all money together' (p<.05).
       There are no statistically significant differences by demographic group among cohabiting couples. 
       Additional details about data access, variable construction, cell sizes and distribution, and chi-squared tests 
       between and within marital groups are available at: https:&#47;&#47;github.com&#47;jrpepin&#47;NCHAT_Money.") 
p1
```


```{r fig-1, echo=FALSE, eval = FALSE}
# save the image as png file
ggsave(filename = file.path(projDir, "waffle.png"), p1, width=6.5, height=6.5, bg = "white", units="in", dpi=300)
```
  
  
<p>  
<p>  
  
Increases in couples more often cohabiting before marrying, attitude data (@pepin_beliefs_2019), and behavioral estimates from outside the U.S. (@hu_divergent_2021) and from non-representative U.S. samples (@bank_of_america_better_2018) all suggest younger couples may be more likely than their older counterparts to keep their finances separate. Yet, this data visualization shows that money arrangements look remarkably similar across age groups. About two-thirds of married couples pool their money, regardless of age, compared with less than one-quarter of cohabitors (p<.05).    
  
Overall, the figure shows that U.S. couples’ money arrangements remain starkly differentiated by marital status and cohort differences are less apparent than expected based on empirical data from other contexts and composition changes among U.S. families that are associated with a reduction in pooled finances. These bivariate statistics open many possibilities for use of NCHAT data to answer further questions related to U.S. couples’ treatment of money that have so far remained elusive.   
  
  
# References

