---
title: "Supplementary Material"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
bibliography: references.json
---
  
```{r setup, echo = FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

## WARNING: Remove the leading # to install packages below first time. 
## Change filepaths below first time

  # if (!require(remotes)) install.packages("remotes") remotes::install_github("fsolt/icpsrdata") # download ICPSR data
  # devtools::install_git("https://git.rud.is/hrbrmstr/waffle.git")                               # install waffle package not on CRAN
  # install.packages("pacman")                                                                    # Install pacman package if not installed

# Installs and loads packages automatically
library("pacman")                  # Load pacman package

# Install packages not yet installed & load them
pacman::p_load(
       here,       # relative file paths
       devtools,   # loading github packages
       foreign,    # read data
       dplyr,      # variable processing
       forcats,    # reverse factor variables
       srvyr,      # calc % with survey weights
       MESS,       # round prop & preserve sum to 100%
       gtsummary,  # pretty weighted tables
       ggplot2,    # graphing
       colorspace) # color palettes   

library("waffle")                 
library("icpsrdata")

```
  
  
```{r packages, echo = FALSE}
# Set the project directory to the current working directory.
# Change the filepath to where the data was downloaded.
projDir <- here::here()             # File path to this project's directory
dataDir <- "./../../Data/NCHAT"     # File path to where data will be downloaded

```
  
  
# Data  
This visualization uses population-based U.S. data from the 2020–2021 [National Couples' Health and Time Study](https://www.icpsr.umich.edu/web/DSDR/studies/38417). The Eunice Kennedy Shriver National Institute of Child Health and Human Development (NICHD) supported the data collection of 3,639 partnered individuals through the Gallup Panel, including an over sample of sexual and gender diverse families. When weighted, the survey provides nationally representative data on cohabiting and married individuals (aged 20-60).  
  
```{r data-1, echo=FALSE, eval = FALSE}
# The icpsr_download function will ask for the login information required by ICPSR: the user’s email and password.
## URL for manual download: https://www.icpsr.umich.edu/web/DSDR/studies/38417
icpsr_download( file_id = 38417,
                download_dir = file.path(dataDir))
```
  
```{r data-2, echo=FALSE, results= 'hide'}
# Load the data and create a new dataframe containing only the variables of interest.  

load(file=file.path(dataDir, "ICPSR_38417/DS0001", '38417-0001-Data.rda'))

data <- da38417.0001 %>%
  select(RESPONSEID, WEIGHT_MAINRESPONDENT, 
         Q40,       # money treatment
         MARITAL_1, # legal marital status
         C18D_B,    # parent identifier using skip logic
                    # {ends_with("HHR16"), starts_with("A1_HHR17_"), starts_with("A1_HHR18_") is masked by ICPSR]}
         D2, HHR5,  # couple type
         AGE_SD)
```
  
  
# Measures  
```{r vars-1, echo=FALSE}
### creates a function so level order remains the same as argument order
fct_case_when <- function(...) {
  args <- as.list(match.call())
  levels <- sapply(args[-1], function(f) f[[3]])  # extract RHS of formula
  levels <- levels[!is.na(levels)]
  factor(dplyr::case_when(...), levels=levels)
}
```
  
  
## _Money arrangement_  
Respondents were told: "Couples handle their money differently. Which of the following do you do?" The response choices were: (1) We keep all of our money separately, (2) Put some money together, or (3) Put all of our money together.  Only 6 (.2%) respondents refused to answer the question item.  

##  _Marital status_
This visualization uses legal marriage status to differentiate married and cohabiting couples. Respondents who answered affirmatively when asked "Have you and your current spouse/partner ever experienced any of the following with each other? Legal marriage" were considered legally married. Four respondents refused to answer the question.   
  
## _Parental status_  
The direct parent identifiers are only available through a restricted Virtual Data Enclave, requiring institutional IRB approval and an application. With reproducibility a priority, parents were instead identified using the skip logic of a publicly available variable on respondents’ stress over dividing childcare with their partner. Respondents who were not shown the question (C18D_B) were determined to be not-parenting and those shown the question were coded as parenting.  
  
## _Couple type_  
Two variables were used to identify the couple type: D2 (Which of the following best describes your gender?) and HHR5 (How would [Field-HHP1] describe their gender?). Women who reported their partner was a man (and vice versa) were coded as different-gender couples. Men who reported their partner was a man and women who reported their partner was a woman were coded as same-gender couples. Respondents who chose another gender for themselves or their partners were not included in this data visualization.
  
## _Age group_
The survey contained a variable (AGE_SD) indicating the respondents' age was between one of three ranges: (1) 18-34, (2) 35-50, or (3) 51-69. The codebook and another variable (AGE_SD_2CAT) indicate the possible age range for main respondents is 20-60. Therefore, the categories were recoded to reflect the true top/bottom ages (partner's age may be younger/older). The refusal rate for this item was 1.6% (n = 57).  
  
```{r vars-2}

## Process variables
data <- data %>%
  mutate(
    # money treatment
    money = fct_case_when(
      Q40 == "(3) Put all of our money together"     ~ "Put all money together",
      Q40 == "(2) Put some money together"           ~ "Put some money together",
      Q40 == "(1) We keep all of our money separate" ~ "Keep all money separate"),
    # marital status
    marco = fct_case_when(
      MARITAL_1  == "(1) Legal marriage"             ~ "Married",
      TRUE                                           ~ "Cohabiting"),
    # parental status
    parent = fct_case_when(
      !is.na(C18D_B)                                 ~ "Parenting",
      TRUE                                           ~ "Not parenting"),
    # couple type
    couple = fct_case_when(
     (D2 == "(1) Man"   & HHR5 == "(2) Woman")       |
     (D2 == "(2) Woman" & HHR5 == "(1) Man")         ~ "Different-gender",
      D2 == "(2) Woman" & HHR5 == "(2) Woman"        ~ "Woman-woman",
      D2 == "(1) Man"   & HHR5 == "(1) Man"          ~ "Man-man"),
    # age
    age = fct_case_when(
      AGE_SD == "(1) 18-34"                          ~ "Age: 20-34", # change bottom code to match codebook
      AGE_SD == "(2) 35-50"                          ~ "Age: 35-50",
      AGE_SD == "(3) 51-69"                          ~ "Age: 51-60")) # change top code to match codebook
```
  
  
# Sample  

This data visualization is based on same- and different-gender U.S. couples aged 20–60 who had no missing information regarding their money arrangements, marital and parental status, and age group. The final sample included 3,537 partnered individuals. More details about the sample recruitment can be found [online](https://osf.io/preprints/socarxiv/p8a34/).  
  
```{r survey}

waffle_svy <- data %>%
  select(money, marco, parent, couple, age, WEIGHT_MAINRESPONDENT, RESPONSEID) %>%
  drop_na() %>% # drop respondents with any missing data
  as_survey_design(id = RESPONSEID,
                   weights = WEIGHT_MAINRESPONDENT)

waffle_svy %>%
  select(c(-RESPONSEID, -WEIGHT_MAINRESPONDENT)) %>%
  tbl_svysummary(statistic = all_categorical() ~ "{n_unweighted} ({p}%)",
                 label = list(money    ~ "Money Treatment",
                              marco    ~ "Marital Status",
                              parent   ~ "Parental Status",
                              couple   ~ "Couple Type",
                              age      ~ "Age"))
 
```
  
<p>  
<p>    
  
  
The bivariate statistics were constructed using the survey-provided variable (WEIGHT_MAINRESPONDENT) to correct for over- or under-sampling of specific target populations. 
  
```{r waffle-1, results='hide'}

# all -- just for reference
df0 <- waffle_svy %>%
  group_by( money) %>%
  # create summary proportions
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>% 
  # round with preserving to 100%
  mutate(vals     = MESS::round_percent(vals), 
         vals_low = MESS::round_percent(vals_low), 
         vals_upp = MESS::round_percent(vals_upp))

# marital status
df1 <- waffle_svy %>%
  group_by(marco, money) %>%
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>% 
  mutate(vals     = MESS::round_percent(vals),
         vals_low = MESS::round_percent(vals_low), 
         vals_upp = MESS::round_percent(vals_upp), 
         group    = "parent",
         subgroup = "All")

# parents
df2 <- waffle_svy %>%
  group_by(marco,parent, money) %>%
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>%
  mutate(vals     = MESS::round_percent(vals), 
         group    = "parent") %>%
  rename(subgroup = parent)

# couple type
df3 <- waffle_svy %>%
  group_by(marco,couple, money) %>%
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>%
  mutate(vals     = MESS::round_percent(vals), 
         group    = "couple") %>%
  rename(subgroup = couple)

# age groups
df4 <- waffle_svy %>%
  group_by(marco,age, money) %>%
  summarize(vals  = survey_mean(na.rm = TRUE, vartype = "ci")) %>%
  mutate(vals     = MESS::round_percent(vals), 
         group    = "age") %>%
  rename(subgroup = age)

# Combine datasets
waffle_data <- df1 %>%
  full_join(df2) %>%
  full_join(df3) %>%
  full_join(df4) %>%
    mutate(group = as_factor(group),
    subgroup = as_factor(subgroup))

```
  
Replication code for data access, variable construction, and data visualization creation is available at [github repository redacted for peer review].  